---
title: "Data Journalism 2"
author: "Derek Nakagawa"
date: "3/29/2022"
---

Importing the data sets from https://sports-statistics.com/sports-data/soccer-datasets/ and combining them so it is easier to use.  In order to combine the files, we needed to make sure they all had the same columns so I used the select function to take the important columns and ignore the columns we don't need (like the bettings odds columns).

```{r setup, include=FALSE, echo=FALSE}
library(stringr)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(stringi)

Liga1 <- read.csv(file = "Liga1-1998-99.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR)
Liga2 <- read.csv(file = "Liga1-1999-00.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR)
Liga3 <- read.csv(file = "Liga1-2000-01.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR)
Liga4 <- read.csv(file = "Liga1-2001-02.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR)
Liga5 <- read.csv(file = "Liga1-2002-03.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR)
Liga6 <- read.csv(file = "Liga1-2003-04.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR)
Liga7 <- read.csv(file = "Liga1-2004-05.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR)

SerieA1 <- read.csv(file = "SerieA-2004-05.csv")
SerieA2 <- read.csv(file = "SerieA-2005-06.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
SerieA3 <- read.csv(file = "SerieA-2006-07.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
SerieA4 <- read.csv(file = "SerieA-2007-08.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
SerieA5 <- read.csv(file = "SerieA-2008-09.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
SerieA6 <- read.csv(file = "SerieA-2009-10.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
SerieA7 <- read.csv(file = "SerieA-2010-11.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)

LaLiga1 <- read.csv(file = "LaLiga-2009-10.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR) 
LaLiga2 <- read.csv(file = "LaLiga-2010-11.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
LaLiga3 <- read.csv(file = "LaLiga-2011-12.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
LaLiga4 <- read.csv(file = "LaLiga-2012-13.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
LaLiga5 <- read.csv(file = "LaLiga-2013-14.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
LaLiga6 <- read.csv(file = "LaLiga-2014-15.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)

Prem1 <- read.csv(file = "PL-2000-01.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
Prem2 <- read.csv(file = "PL-2001-02.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
Prem3 <- read.csv(file = "PL-2002-03.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
Prem4 <- read.csv(file = "PL-2003-04.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
Prem5 <- read.csv(file = "PL-2004-05.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
Prem6 <- read.csv(file = "PL-2005-06.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
Prem7 <- read.csv(file = "PL-2006-07.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
Prem8 <- read.csv(file = "PL-2007-08.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)

Prem9 <- read.csv(file = "PL-2012-13.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
Prem10 <- read.csv(file = "PL-2013-14.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
Prem11 <- read.csv(file = "PL-2014-15.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
Prem12 <- read.csv(file = "PL-2015-16.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
Prem13 <- read.csv(file = "PL-2016-17.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
Prem14 <- read.csv(file = "PL-2017-18.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
Prem15 <- read.csv(file = "PL-2018-19.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)
Prem16 <- read.csv(file = "PL-2019-20.csv") %>% select(Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HS, AS, HST, AST, HC, AC, HF, AF, HY, AY, HR, AR)


Porto <- rbind(Liga1, Liga2, Liga3, Liga4, Liga5, Liga6, Liga7)
Porto <- na.omit(Porto)
Porto <- Porto %>% separate(Date, c("Day", "Month", "Year")) %>% filter(HomeTeam == 'Porto' | AwayTeam == 'Porto')
Porto$Date <- paste(stri_sub(Porto$Year, -2), "/", Porto$Month, "/", Porto$Day) 
Porto <- subset(Porto, select = -c(Day:Year))
Porto$Team <- "Porto"

Chelsea1 <- rbind(Prem1, Prem2, Prem3, Prem4, Prem5, Prem6, Prem7, Prem8)
Chelsea1 <- na.omit(Chelsea1)
Chelsea1 <- Chelsea1 %>% filter(HomeTeam == 'Chelsea' | AwayTeam == 'Chelsea')
Chelsea1$Team <- "Chelsea1"

InterMilan <- rbind(SerieA2, SerieA3, SerieA4, SerieA5, SerieA6, SerieA7)
InterMilan <- na.omit(InterMilan)
InterMilan <- InterMilan %>% filter(HomeTeam == 'Inter' | AwayTeam == 'Inter')
InterMilan$Team <- "Inter Milan"

RealMadrid <- rbind(LaLiga1, LaLiga2, LaLiga3, LaLiga4, LaLiga5, LaLiga6)
RealMadrid <- na.omit(RealMadrid)
RealMadrid <- RealMadrid %>% filter(HomeTeam == 'Real Madrid' | AwayTeam == 'Real Madrid')
RealMadrid$Team <- "Real Madrid"

Chelsea2 <- rbind(Prem9, Prem10, Prem11, Prem12)
Chelsea2 <- na.omit(Chelsea2)
Chelsea2 <- Chelsea2 %>% filter(HomeTeam == 'Chelsea' | AwayTeam == 'Chelsea')
Chelsea2$Team <- "Chelsea2"

ManUnited <- rbind(Prem11, Prem12, Prem13, Prem14, Prem15, Prem16)
ManUnited <- na.omit(ManUnited)
ManUnited <- ManUnited %>% filter(HomeTeam == 'Man United' | AwayTeam == 'Man United')
ManUnited$Team <- "Manchester United"

All_Teams <- rbind(Chelsea1, InterMilan, RealMadrid, Chelsea2, ManUnited) %>%
  separate(Date, c("Day", "Month", "Year"))

All_Teams$Date <- paste(stri_sub(All_Teams$Year, -2), "/", All_Teams$Month, "/", All_Teams$Day) 
All_Teams <- subset(All_Teams, select = -c(Day:Year))

```


Need to clean the data so it can be used for graphs and comparing the managers



```{r setup, include=FALSE, echo=FALSE}
Final_DataSet <- data.frame(
  Team_Order = c("1. Porto", "1. Porto", "1. Porto", "2. Chelsea1", "2. Chelsea1", "2. Chelsea1", "3. Inter Milan", "3. Inter Milan", "3. Inter Milan", "4. Real Madrid", "4. Real Madrid", "4. Real Madrid", "5. Chelsea2", "5. Chelsea2", "5. Chelsea2", "6. Man United", "6. Man United", "6. Man United"),
  Team = c("Porto", "Porto", "Porto", "Chelsea1", "Chelsea1", "Chelsea1", "Inter Milan", "Inter Milan", "Inter Milan", "Real Madrid", "Real Madrid", "Real Madrid", "Chelsea2", "Chelsea2", "Chelsea2", "Manchester United", "Manchester United", "Manchester United"),
  Team_Name_Dataset = c("Porto", "Porto", "Porto", "Chelsea", "Chelsea", "Chelsea", "Inter", "Inter", "Inter", "Real Madrid", "Real Madrid", "Real Madrid", "Chelsea", "Chelsea", "Chelsea", "Man United", "Man United", "Man United"),
  Manager = c("1 Previous Manager", "2 Jose Mourinho", "3 Subsequent Manager"),
  Games = 0,
  Wins = 0,
  Loses = 0,
  Ties = 0,
  WinPercent = 0,
  
  Goals_PG = 0,
  Goals_Against_PG = 0,
  Shots_PG = 0,
  Shots_Against_PG = 0, 
  Shots_On_Goal = 0,
  Shots_On_Goal_Against = 0,
  
  Fouls_PG = 0, 
  Num_Y_Cards_PG = 0,
  Num_R_Cards_PG = 0
)

manager_dates <- c("00 / 09 / 17", "04 / 05 / 31", "04 / 06 / 02", "07 / 09 / 20", "07 / 09 / 19", "08 / 05 / 24", 
                   "04 / 07 / 01", "08 / 05 / 30", "08 / 06 / 02", "10 / 05 / 28", "10 / 06 / 10", "10 / 12 / 23", 
                   "09 / 06 / 02", "10 / 05 / 26", "10 / 05 / 31", "13 / 06 / 01", "13 / 06 / 25", "15 / 05 / 25", 
                   "12 / 11 / 21", "13 / 05 / 27", "13 / 06 / 03", "15 / 12 / 17", "15 / 13 / 09", "16 / 06 / 30", 
                   "14 / 07 / 16", "16 / 05 / 23", "16 / 05 / 27", "18 / 12 / 17", "18 / 12 / 19", "21 / 11 / 21")


for(i in 4:nrow(Final_DataSet)){

  date_num <- seq(1, 100, 2)
  Date1 <- manager_dates[date_num[i-3]]
  Date2 <- manager_dates[date_num[i-3] + 1]
  
  Team_New <- All_Teams %>%
    filter(Team == Final_DataSet$Team[i]) %>%
    filter(Date >= Date1 & Date <= Date2)
  Home_Team_New <- Team_New %>%
    filter(HomeTeam == Final_DataSet$Team_Name_Dataset[i])
  Away_Team_New <- Team_New %>%
    filter(AwayTeam == Final_DataSet$Team_Name_Dataset[i])
  
  Games = nrow(Team_New)
  Final_DataSet$Games[i] = nrow(Team_New)
  Final_DataSet$Goals_PG[i] = ( sum(Home_Team_New$FTHG) + sum(Away_Team_New$FTAG) ) / Games
  Final_DataSet$Goals_Against_PG[i] = ( sum(Home_Team_New$FTAG) + sum(Away_Team_New$FTHG) ) / Games
  Final_DataSet$Shots_PG[i] = ( sum(Home_Team_New$HS) + sum(Away_Team_New$AS) ) / Games
  Final_DataSet$Shots_Against_PG[i] = ( sum(Home_Team_New$AS) + sum(Away_Team_New$HS) ) / Games
  Final_DataSet$Shots_On_Goal[i] = ( sum(Home_Team_New$HST) + sum(Away_Team_New$AST) ) / Games
  Final_DataSet$Shots_On_Goal_Against[i] = ( sum(Home_Team_New$AST) + sum(Away_Team_New$HST) ) / Games
  Final_DataSet$Fouls_PG[i] = ( sum(Home_Team_New$HF) + sum(Away_Team_New$AF) ) / Games
  Final_DataSet$Num_Y_Cards_PG[i] = ( sum(Home_Team_New$HY) + sum(Away_Team_New$AY) ) / Games
  Final_DataSet$Num_R_Cards_PG[i] = ( sum(Home_Team_New$HR) + sum(Away_Team_New$AR) ) / Games
    
  Home_Team_Win <- Home_Team_New %>%
    filter(FTR == "H")
  Away_Team_Win <- Away_Team_New %>%
    filter(FTR == "A")
  Final_DataSet$Wins[i] = nrow(Home_Team_Win) + nrow(Away_Team_Win)
  
  Home_Team_Loss <- Home_Team_New %>%
    filter(FTR == "A")
  Away_Team_Loss <- Away_Team_New %>%
    filter(FTR == "H")
  Final_DataSet$Loses[i] = nrow(Home_Team_Loss) + nrow(Away_Team_Loss)
  
  Final_DataSet$Ties[i] = Games - Final_DataSet$Loses[i] - Final_DataSet$Wins[i]
  Final_DataSet$WinPercent[i] = Final_DataSet$Wins[i] / Games
}



# Fill in data for just Porto Statistics
#----------------------------------------------------------------------------------

Porto_Previous <- Porto %>%
    filter((Date >= "98 / 06 / 04" & Date <= "99 / 12 / 31") | (Date >= "00 / 01 / 01" & Date <= "01 / 06 / 08"))

Home_Team_New1 <- Porto_Previous %>%
    filter(HomeTeam == "Porto")
Away_Team_New1 <- Porto_Previous %>%
    filter(AwayTeam == "Porto")

Home_Team_Win1 <- Home_Team_New1 %>%
    filter(FTR == "H")
Away_Team_Win1 <- Away_Team_New1 %>%
    filter(FTR == "A")
Final_DataSet$Wins[1] = nrow(Home_Team_Win1) + nrow(Away_Team_Win1)

  
Home_Team_Loss1 <- Home_Team_New1 %>%
    filter(FTR == "A")
Away_Team_Loss1 <- Away_Team_New1 %>%
    filter(FTR == "H")
Final_DataSet$Loses[1] = nrow(Home_Team_Loss1) + nrow(Away_Team_Loss1)

Games <- nrow(Porto_Previous)

Final_DataSet$Goals_PG[1] = ( sum(Home_Team_New1$FTHG) + sum(Away_Team_New1$FTAG) ) / Games
Final_DataSet$Goals_Against_PG[1] = ( sum(Home_Team_New1$FTAG) + sum(Away_Team_New1$FTHG) ) / Games
  
Final_DataSet$Games[1] = nrow(Porto_Previous)
Final_DataSet$Ties[1] = Games - Final_DataSet$Loses[1] - Final_DataSet$Wins[1]
Final_DataSet$WinPercent[1] = Final_DataSet$Wins[1] / Games



Porto_Manager_Dates = c("02 / 01 / 23", "04 / 06 / 02", "04 / 08 / 11", "05 / 01 / 31")
for(i in 2:3){

  date_num <- seq(1, 100, 2)
  Date1 <- Porto_Manager_Dates[date_num[i-1]]
  Date2 <- Porto_Manager_Dates[date_num[i-1] + 1]
  
  Team_New <- Porto %>%
    filter(Date >= Date1 & Date <= Date2)
  Home_Team_New <- Team_New %>%
    filter(HomeTeam == "Porto")
  Away_Team_New <- Team_New %>%
    filter(AwayTeam == "Porto")

  Home_Team_Win <- Home_Team_New %>%
    filter(FTR == "H")
  Away_Team_Win <- Away_Team_New %>%
    filter(FTR == "A")
  Final_DataSet$Wins[i] = nrow(Home_Team_Win) + nrow(Away_Team_Win)
  
  Home_Team_Loss <- Home_Team_New %>%
    filter(FTR == "A")
  Away_Team_Loss <- Away_Team_New %>%
    filter(FTR == "H")
  Final_DataSet$Loses[i] = nrow(Home_Team_Loss) + nrow(Away_Team_Loss)
  
  Games = nrow(Team_New)
  Final_DataSet$Games[i] = nrow(Team_New)
  
  Final_DataSet$Goals_PG[i] = ( sum(Home_Team_New$FTHG) + sum(Away_Team_New$FTAG) ) / Games
  Final_DataSet$Goals_Against_PG[i] = ( sum(Home_Team_New$FTAG) + sum(Away_Team_New$FTHG) ) / Games
  Final_DataSet$Ties[i] = Games - Final_DataSet$Loses[i] - Final_DataSet$Wins[i]
  Final_DataSet$WinPercent[i] = Final_DataSet$Wins[i] / Games
}
 
Final_DataSet$Shot_Difference <- Final_DataSet$Shots_On_Goal - Final_DataSet$Shots_On_Goal_Against

```


Final Datasheets are now finished


```{r setup}
Final_DataSet

Final_DataSet_No_Porto <- Final_DataSet[-c(1:3), ]
Final_DataSet_No_Porto
```


Things to look at: WinPercent, Games, Goals_PG, Goals_Against_PG, Shots_On_Goal, Shots_On_Goal_Against


```{r setup, include=FALSE, echo=FALSE}

# win percentages
ggplot(Final_DataSet, aes(x=Team_Order, y=WinPercent, fill=Manager)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("azure4", "red", "darkgray"), labels = c("Previous Manager", "Jose Mourinho", "Subsequent Manager")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(label = Games), vjust = 1.5, position = position_dodge(.9), size = 7) +
  geom_text(aes(label = "Games"), vjust = 4.5, position = position_dodge(.9), size = 4) +
  theme(axis.title = element_text(size = 30), axis.text = element_text(size = 20), plot.title = element_text(size = 30), legend.title = element_text(size = 20), legend.text = element_text(size = 15)) +
  labs(
    x = "Soccer Team",
    y = "Win Percentage",
    title = "Win Percentages for Managers in Regular Season",
    fill = "Manager"
  )


#shots on Goal
ggplot(Final_DataSet_No_Porto, aes(x=Team_Order, y=Shot_Difference, fill=Manager)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("azure4", "red", "darkgray"), labels = c("Previous Manager", "Jose Mourinho", "Subsequent Manager")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Soccer Team",
    y = "Average Shots on Goal Per Game",
    title = "Shots on Goal Per Game for Managers in Regular Season",
    fill = "Manager"
  )


#shots on goal against
ggplot(Final_DataSet_No_Porto, aes(x=Team_Order, y=Goals_PG, fill=Manager)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("azure4", "red", "darkgray"), labels = c("Previous Manager", "Jose Mourinho", "Subsequent Manager")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Soccer Team",
    y = "Average Shots on Goal Against Per Game",
    title = "Goals Against Per Game for Managers in Regular Season",
    fill = "Manager"
  )

Porto
```






Look at the placing of his teams

```{r setup, include=FALSE, echo=FALSE}
library(ggplot2)
library(png)
library(grid)
library(ggimage)

seasons_position <- data.frame(
  Season = c("01/02", "02/03", "03/04", 
             "04/05", "05/06", "06/07", 
             "08/09", "09/10", 
             "10/11", "11/12", "12/13", 
             "13/14", "14/15", 
             "16/17", "17/18"),
  Team = c("Porto", "Porto", "Porto", 
           "Chelsea", "Chelsea", "Chelsea", 
           "Inter Milan", "Inter Milan", 
           "Real Madrid", "Real Madrid", "Real Madrid", 
           "Chelsea", "Chelsea", 
           "Man United", "Man United"),
  place_in_league = c(3, 1, 1, 
                      1, 1, 2, 
                      1, 1, 
                      2, 1, 2,
                      3, 1,
                      6, 2),
  image = c("Porto.png", "Porto.png", "Porto.png",
            "Chelsea.png", "Chelsea.png", "Chelsea.png", 
            "Inter_Milan.png", "Inter_Milan.png", 
            "Real_Madrid.png", "Real_Madrid.png", "Real_Madrid.png", 
            "Chelsea.png", "Chelsea.png", 
            "Man_United.png", "Man_United.png")
)


```


```{r setup}
ggplot(data = seasons_position, aes(x=Season, y=place_in_league, group = 1)) +
  geom_line(color = "red", size = 1.5) +
  geom_image(aes(image = image), size = 0.06, asp = 1.6) + 
  expand_limits(x = 0, y = 1) +
  theme(axis.title = element_text(size = 20), axis.text = element_text(size = 15), plot.title = element_text(size = 30)) +
  labs(
    x = "Season Managed",
    y = "Placement in League (Lower the Better)",
    title = "Jose Mourinho Placement in League",
  )

```



